# pml data analysis
# Author: Alexander Zharov


## Reading data:

```{r}
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")
```

## Cleaning data:

Now let's make a summary of data. Output will be omitted for the sake of brevity.

There are a lot of columns in a dataset, consisting of a lot of NA values. Let's exclude them from dataset. New dataset will contain 93 of 160 original dataset columns.

```{r, results=FALSE}
train_summary <- summary(training)
columns <- grepl("NA's", train_summary[7, ])
train <- training[ , !columns]
test <- testing[ , !columns]
```

The same way let's exclude column which mostly doesn't contain values. New dataset will contain 59 of 160 original dataset columns.

```{r, results=FALSE}
train_summary <- summary(train)
columns <- grepl(":19216", train_summary[1, ])
train <- train[ , !columns]
test <- test[ , !columns]
```

Now let's exclude variables that evidently can't impact doing exersize. First, this is the number of observation.
Dataset is created artificially, so it doesn't necessarily have time pattern of real world observations. So, time variables are also excluded.

```{r}
train <- train[ , -1 * c(1, 3, 4, 5, 6)]
test <- test[ , -1 * c(1, 3, 4, 5, 6, 59)]
```

## Splitting data:

Let's divide training data into 2 parts - train and valid data.

```{r}
library(caret)
set.seed(1234)
inTrain <- createDataPartition(y=train$classe, p=0.75, list=FALSE)
train2 <- train[inTrain, ]
valid2 <- train[-inTrain, ]
```

## Fitting model:

Now fit model with bagging on train dataset...

```{r, results=FALSE, warning=FALSE}
library(ipred)
modFit <- train(classe ~ ., data = train2, method = "treebag")
```

... and test its accuracy on validation dataset

```{r}
pred <- predict(modFit, valid2)
valid2$predRight <- pred == valid2$classe
table(pred, valid2$classe)
```

The correctly identified classes percentage in validation dataset is:
```{r}
sum(diag(table(pred, valid2$classe))) * 100 / nrow(valid2)
```

Not bad!

## Apply to test:

Let's apply our model to test data 20 observations for the 2nd project:
```{r}
answers <- predict(modFit, test)
answers
```